Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset

### 2025-03-01 code/manuscript.Rmd
push for second draft review

---
title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large Open Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods relying on pre- and post-  retrofit comparison for estimating whole-building energy savings are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy occur during the M&V process. Those events are unrelated to the intervention but significantgly affect building energy consumption and thus when the analyst applies conventiona M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from random sampling, where the analyst randomly selects whether to implement the baseline or the intervention strategy each day. We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that comparing to the conventional method, the randomized method provides faster and more accurate savings estimation. Additionaly we found that the randomized M&V approach estimates much closer to the true savings when non-routine events are present demonstrating much improved reliability.

@@

## Background

### Conventional M&V

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008,
@efficiency_valuation_organisation_international_2007,
@ashrae_ashrae_2014]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters. The corresponding M&V process required
by ASHRAE Guideline 14, which we refer to as the convention method in
this paper. Typically, the process begins with baseline measurements
taken over a year before implementing any energy-efficiency retrofit,
followed by the same measurement procedure during the intervention
period. After collecting two years of data, an M&V analyst fits an
energy prediction model, using variables such as outdoor temperature and
time [@mathieu_quantifying_2011] to project baseline energy consumption
in the post-retrofit period. The difference between the counter-factual
baseline and the measured intervention represents the energy savings. A
key drawback of this method is its reliance on a two-year timeline to
quantify savings, during which baseline measurements can become outdated
due to changes in building performance caused by non-routine events
unrelated to the intervention. This limitation reduces the feasibility
of rapid M&V and complicates the quantification of estimation
uncertainty, thus impacting the accuracy and timeliness of savings
assessments.

### Randomized M&V

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we propose a novel M&V method
that adopts the randomized crossover design, a concept borrowed from
medical and agricultural studies. Another improvement is that we
proposed a sequential evaluation framework and defined stopping criteria
to end the M&V if the target effect is detected. This is to avoid
unnecessary measurement collection over the full 2-year M&V cycle. The
full framework is detailed in another study with all stopping criteria
outlined <!--# TODO: Add references -->. In summary, this method
provides M&V analysts with a randomized schedule that alternates between
baseline and intervention implementation while ensuring balanced
sampling across days of the week. For example, given a 10-week M&V
period for 1 intervention, the balanced randomized schedule would
equally sample 5 Mondays with the baseline and 5 Mondays with the
intervention. The limitation of the randomized M&V is that it is only
applicable to a subset of retrofit projects such as control
interventions. However, for all applicable use cases, it allows analysts
to detect energy savings sequentially as the study progresses meaning
once the desired savings target is achieved, analysts can terminate the
M&V. The key advantage of randomization, which is one of the study
objectives, is that if control strategies are sampled with equal
probability, the influence of non-routine events is likely to be evenly
distributed between the baseline and intervention measurements. This
means that the savings estimate, calculated as the difference between
the two, effectively 'cancels out' the impact of these disturbances,
leading to a more accurate and unbiased assessment of the intervention's
impact.

### BDG2 dataset

The Building Genome Dataset 2 (BGD2) is an extensive open-access dataset
designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 500 buildings' metadata and realistic operational
information from across North America and Europe, making it one of the
most comprehensive collections of building-related data available for
scientific use. The dataset includes various commercial building types
such as offices, education facilities, public, and retail buildings, and
provides detailed information on their physical characteristics (e.g.
enregy ratings, heating types and floor area) and hourly measurements of
chilled and hot water, electricity, gas usage as well as site outdoor
weather conditions.

@@

Most research related to M&V for whole-building approach focuses on the
accuracy of baseline modeling, exploring model performance from simple
regression models to more complex machine learning techniques. One study
reviewed various models suitable for M&V applications as well as
selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015,
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014, @heo_calibration_2012,
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that baseline model might deteriorate (i.e.
becomes 'stale') over an extended period of pre- and post-analysis.

@@

A common non-routine event in energy-saving M&V projects is a change in
occupancy or a significant shift in occupant behavior. These changes can
greatly affect measured energy consumption in buildings and are
typically unrelated to the intervention strategy. For instance, during
the COVID-19 pandemic in 2020, most commercial buildings were
unoccupied, leading to a noticeable drop in energy bills despite no
energy-efficiency measures being implemented [@cai_impact_2024,
@chihib_impact_2021, @gaspar_assessing_2022, @kang_changes_2021].
Furthermore, subsequent research has shown that hybrid working modes,
allowing employees to work remotely, have persisted after the outbreak
of the pandemic [@aksoy_working_2022], adding further complexity to
energy consumption patterns due to evolved occupant behaviors with
higher flexibility [@gui_impact_2021, @mantesi_office_2022,
@xie_does_2021]. One study realized the limitation of current M&V
methods, which only consider adjusting for outdoor weather, is
insufficient and emphasized the importance of requiring matched
comparison groups to control for exogenous factors beyond weather
differences when comparing between baseline and intervention
[@demand_side_analytics_population_2022]. Another type of non-routin
event is filter clogging in air handling units due to particle
accumulation. This can cause supply fans to gradually consume more
energy to maintain required duct static pressure [@feng_newly_2019,
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be underestimated as
increased energy use is incorrectly attributed to the intervention
rather than the mechanical issue.

@@

As mentioned, the goal of an M&V project is to determine the
effect---typically energy savings---of an energy-efficient intervention.
In this study, we focus primarily on switchable interventions, which
often involve control retrofits. An example of such an intervention is a
control retrofit developed by a software-as-a-service company that
adjusts the chilled water plant's supply water temperature based on
outdoor weather conditions [@roa_field_2023,
@lee_simulationoptimization_2012, @qiu_chilled_2022]. Therefore, we
defined the M&V scenario as follows:

*"A company aims to sell its supply temperature reset control software
package to a customer, such as a building owner, with a guarantee that
it will reduce the building's electricity usage. If the building owner
decides to purchase the service, the company agrees to charge a service
fee based on a percentage of the measured energy savings."*

As required by the M&V scenario, we assessed the performance of both the
conventional and the novel randomized M&V methods by estimating the
intervention energy savings for all valid buildings in the dataset. By
conducting such analysis, we aim to:

1.  Compare the energy saving estimation accuracy between the
    conventional and the randomized method. This study extends the
    comparison to a large sample of buildings, covering a variety of
    types and climate zones. The comparison metrics include both
    estimation accuracy and M&V finishing timeline.

2.  Verify the enhanced robustness of the randomized method. By using
    realistic measurements from real-world buildings, which include
    various sources of noise, we aim to reflect the challenges faced by
    building analysts in real projects. As will be demonstrated in the
    following sections, the randomized approach is less impacted by
    non-routine events (i.e., measurement 'noise'), resulting in more
    reliable energy savings estimates.

3.  Demonstrate the implementation of the proposed randomized M&V method
    using a public available dataset. We ensured the reproducibility of
    the method by making the analysis code open source including
    randomized schedule generation, sequential statistical analysis,
    energy modeling and normalized saving calculation. Using the
    available open resources, building analysts should be able to
    seamlessly integrate and apply them in their own M&V projects.

@@

```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="50%"}

@@

## Data filtering

@@

4.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water setpoint reset
    control.

```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

@@

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  group_by(type, site) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
  group_by(type, site) %>% 
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "site")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(site, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 10, as.character(n), "")), check_overlap = T) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

For all qualified buildings, we further added one more stringent
criterion to label all buildings with very stable electricity usage
between the two years:

1.  No statistical significant difference (P-value \> 0.05) between the
    two-year electricity usage.

As a results, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption made
by conventional M&V is overly simplified. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events and according to the statistics, is more
commonly observed in real applications.

## Control intervention implementation

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022, @congradac_recognition_2012]. For both
strategies, we assume that the chiller is activated when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

@@

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5.5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Intervention logic: chilled water supply temperature reset") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building, and were used uniformly across the dataset.

|                        |                                                                  |       |
|------------------------|------------------------------------------------------------------|-------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

## M&V methods comparison

We described more in detail the workflow of both conventional and
randomized M&V methos in the previous study [@RAFTERY2024111134].

In a similar style, we summarized the experimental design as:

-   The HVAC system operates from 06:00 to 22:00 each day, so we use a
    daily sampling interval with the sampling time at midnight each day.

@@

and stopping criteria:

@@

-   At least 80% of the drybulb temperature range in the annual TMY data
    sampled by both strategies.

@@

-   As no baseline data is available, test with an equal sampling ratio
    (50% baseline, 50% intervention).

@@

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=10}

@@

@@

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=9, fig.width=12}

>>>>

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy, or equipment failures occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy each day. We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimation. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.

@@

## Conventional M&V

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. Typically, the process begins with
baseline measurements taken during the year before implementing the 
energy-efficiency retrofit, followed by the same measurement procedure
during the year after the intervention. After collecting two years of
data, an M&V analyst fits an energy prediction model, using variables
such as outdoor temperature and time [@mathieu_quantifying_2011] to
project baseline energy consumption in the post-retrofit period. The
difference between the counter-factual baseline and the measured
intervention period energy consumption represents the energy savings.
One of the key drawbacks of this method is its reliance on a two-year
timeline to quantify savings, during which baseline measurements can
become outdated due to changes in building performance caused by
non-routine events unrelated to the intervention. This limitation
reduces the feasibility of rapid M&V and complicates the quantification
of estimation uncertainty, thus impacting the accuracy and timeliness of
savings assessments.

## Randomized M&V

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we propose a novel M&V method
that adopts the randomized crossover design, a gold standard from
medical and agricultural studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. Another improvement is that we
proposed a sequential evaluation framework and defined stopping criteria
to end the M&V if the target effect is detected. This is to avoid
unnecessary measurement collection over the full 2-year M&V cycle. The
full framework is detailed in a previous published study with all
stopping criteria and example usecases outlined [@raftery2024]. In
summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially right after test begins.
In addition, once the desired savings target is achieved, analysts can
terminate the M&V and switch to 100% intervention. The key advantage of
randomization is that if control strategies are sampled with equal
probability, the influence of other confounding factors such as
occupancy change (or commonly known as non-routine events) is likely to
be evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention's impact.

## BDG2 dataset

The Building Genome Dataset 2 (BGD2) is an extensive open-access dataset
designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly use the metered electricity and outdoor weather measurements for
running the M&V methods.

@@

ASHRAE Guidelinge 14 and IPMVP provides options for whole building M&V
use cases. For code compliance using prescriptive option, 12 months of
baseline and 12 months of postretrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline model fitting accuracy should have CV(RMSE) lower than 20%.
Thus, most research related to M&V for whole-building approach focuses
on the accuracy of baseline modeling, exploring model performance from
simple regression models to more complex machine learning techniques.
One study reviewed various models suitable for M&V applications as well
as selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that baseline model might deteriorate (i.e.
becomes 'stale') over an extended period of pre- and post-analysis.

@@

Non-routine events during M&V commonly refers to unexpected changes in a
building that influences its energy usage. These changes can greatly
affect measured energy consumption in buildings and are typically
unrelated to the intervention strategy. Therefore, those changes are
considered 'static factors' respect to measured independent variables. A
common non-routine event in energy-saving M&V projects is a change in
occupancy or a significant shift in occupant behavior, equipment run
time, and operating conditions (e.g. set points, lighting and
ventilation levels). However, current standards or guidelines only
provides general approach for consideration. For example, IPMVP requires
facility owner and the M&V analyst periodically perform inspections of
all equipment and operations during reporting period, which is labor
intensive and error-prone. ASHRAE guideline 14 recommends performing
engineering calculations or computer software simulations to adjust
post-retrofit baseline. Additional, it is very rare to have access to
all measurements such as occupancy, and thus the analyst normally assume
those factors remain unchanged throughout the study. Specifically,
studies investigating the effects of demand response on building energy
efficiency commonly uses linear interpolation to estimate
counter-factual baseline [@beil_round-trip_2015;
@keskar_commercial_2020], and one study points out that it is inaccurate
to assume no change in the operating conditions during the test period
[@huang_experimental_2023]. One likely encountered non-routine event due
to operation condition change is filter clogging in air handling units
due to particle accumulation. This can cause supply fans to gradually
consume more energy to maintain required duct static pressure
[@feng_newly_2019; @zhai_full-scale_2017]. If M&V analysts are unaware
of such changes and lack an appropriate adjustment method (e.g.,
replacing filters before the intervention begins), the savings could be
underestimated as increased energy use is incorrectly attributed to the
intervention rather than the mechanical issue. Some studies realized the
limitation of current M&V methods, which only consider adjusting for
outdoor weather, is insufficient and emphasized the importance of
requiring matched comparison groups to control for exogenous factors
beyond weather differences when comparing between baseline and
intervention [@demand_side_analytics_population_2022;
@huang_experimental_2023]. However, those methods are still unable to
accurately quantify uncertainties due to changing baseline.

@@

As mentioned, the goal of an M&V project is to determine the
effect---typically energy savings---of an switchable energy-efficient
intervention. The goal is to determine, for a large sample of buildings,
how much more accurately the novel M&V method would estimate the savings
of such an intervention compared to conventional M&V, and how much more
quickly it would reach a result. An example of such an intervention is a
control retrofit developed by a software-as-a-service company that
adjusts the chilled water plant's supply water temperature, which can be
commonly found in the literature [@duarte_field_2023;
@lee_simulationoptimization_2012; @qiu_chilled_2022; @jin2007;
@taylor_optimizing_2012]. In our case, we make it even simpler by
adjusting setpoint based on outdoor weather conditions with more
description shown in Section \@ref(apply-virtue-intervention).
Therefore, we defined the M&V scenario as follows:

"A company aims to sell its supply temperature reset control software
package to a customer, in this case, the building owner, with a
promotion that it will reduce the building's electricity usage. If the
building owner decides to purchase the service, the company agrees to
charge a service fee based on a percentage of the measured energy
savings."

As required by the M&V scenario, we assessed the performance of both the
conventional and the novel randomized M&V methods by estimating the
intervention energy savings for all valid buildings in the BG2 dataset.
By conducting such analysis, we aim to:

1.  Compare the energy saving estimation accuracy between the
    conventional and the randomized method. This study extends the
    comparison to a large sample of buildings, covering a variety of
    types and climate zones. The comparison metrics include both
    estimation accuracy and M&V completion timeline.

2.  Verify the enhanced robustness of the randomized method. By using
    realistic measurements from real-world buildings, which include
    various sources of noise, we aim to reflect the challenges faced by
    building analysts in real projects. As we will perform numercial
    simulation on the existing dataset, the ground-truth savings can be
    calculated as the reference for comparison.

3.  Open-source implementation of the proposed randomized M&V method
    using a public dataset. We ensured the reproducibility of the method
    by making the analysis code open source including randomized
    schedule generation, sequential statistical analysis, energy
    modeling and normalized saving calculation. In addition, we also
    included codes for extended use cases of the randomized method, such
    as changing sampling ratio and sampling intervals. Using the
    available open resources, building analysts should be able to
    seamlessly integrate and apply them in their own M&V projects.

@@

```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}

@@

## Filter and clean dataset

@@

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5%
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrenceberkeleynationallab; @mathew_big-data_2015].

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water set point reset
    control.

```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 10 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

@@

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
   left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "location")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(location, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

For all qualified buildings, we further added one more stringent
criterion to explore all buildings with very stable electricity usage
between the two years:

1.  No statistical significant difference (P-value \> 0.05) between the
    two-year electricity usage, which are sometimes assumed to be the
    case for whole-building measurement and verification.

As a results, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption made
by conventional M&V is overly simplified. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events and according to the statistics shown here, is
commonly observed in real applications.

## Apply virtue intervention

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller is activated when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

@@

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building and were applied uniformly in the filtered dataset.
Interested readers can also change the parameters in our open-source
code to simulate different scenarios.

|                        |                                                                  |       |
|--------------|--------------------------------------------|--------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

## Run M&V methods

We described more in detail the workflow of both conventional and
randomized M&V methods in the previous study [@raftery2024]. The
conventional method is a simple pre- and post- comparison on a 12-month
baseline and 12-month intervention timeline. The randomized method
defines sampling requirements as:

-   Use a daily sampling interval with the sampling time at midnight
    each day.

@@

and stopping criteria as:

@@

-   At least 80% of the dry-bulb temperature range in the annual TMY
    data sampled by both strategies.

@@

-   test with an equal sampling ratio (50% baseline, 50% intervention).

@@

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage and $E'$ indicates the
estimated electricity usage either through the conventional method or
the new randomized method. We present the distribution of savings
estimation accuracy using boxen plots, also known as letter plots, which
is an advanced variation of the box plot designed to extend beyond the
interquartile range (IQR) by progressively dividing the data into
smaller percentiles, revealing more detail in the tails of the
distribution. We set the division parameter to k = 4, meaning the entire
box area represents the data distribution from 6.25% to 93.75%. As the
steps move closer to the center line (50% median), the distribution
range progressively narrows with the next step represents 12.5% to
87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
showed the mean value of the distribution on top.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=12}

@@

p_accuracy <- ggarrange(p_top, p_middle, 
                        ncol = 1, nrow = 2,
                        labels = c("a)", "b)"),
                        common.legend = T, 
                        legend = "bottom")

@@

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=6, fig.width=12}
